### 2008-11-04 master-thesis/texto/introducao.tex
- Very first master thesis beginning. There are still some text related to distributed system, because this latex files came from a classmate thesis (thanks, Koslovski o/).

\chapter{Introdução} 
\label{capitulo:introducao}

\section{Contexto e Motivação}

\section{Objetivos e Contribuição}

\section{Organização do Texto}