### 2008-11-17 master-thesis/texto/conclusao.tex
- The skeleton of my master thesis. Someday I believe it will have a lot of text/content and some nice images. That's we hope, we hope...

\chapter{Conclusão e trabalhos futuros} 
\label{capitulo:conclusao}

--Trabalhos futuros:
-- Melhoramento nas funções que geram relevo;
-- Texturização conforme altura do terreno e afins;
-- Rios;
-- GPU;