### 2009-07-23 paper/paper.tex
- More English text (I REALLY don't want to sleep, why should I? Sleep is for weak people :D).

\documentclass[a4paper]{sbgames}               % final
%\usepackage[scaled=.92]{helvet}
\usepackage[latin1]{inputenc}   % pacote para acentuação
\usepackage{times}
\usepackage{graphicx}

%% use this for zero \parindent and non-zero \parskip, intelligently.
\usepackage{parskip}

%% the 'caption' package provides a nicer-looking replacement
\usepackage[labelfont=bf,textfont=it]{caption}

\usepackage{url}

%% Paper title.
\title{Charack: ferramenta para geração em tempo real de mundos virtuais pseudo-infinitos para jogos 3D}

%% Author and Affiliation (multiple authors). Use: and between authors

\author{Fernando Bevilacqua\\ Cesar Tadeu Pozzer \\ Universidade Federal de Santa Maria 
        %\and Name2 B. Surname2\\ Name3 C. Surname3\\ ZZZZ University
        %\and Name4 D. Surname4\\ Farwest Research Center 
}
\contactinfo{\{fernando,pozzer\}@inf.ufsm.br
}
%% Keywords that describe your work.
\keywords{MMO, mundos virtuais, geração de terreno, jogos 3D, ruído, geração procedimental, multifractal}

%% Start of the paper
% Attention: As you need to insert EPS images in Postscript, 
% you need to insert PDF images into PDFs. 
% In the text, extensions cancbe omitted (latex use .eps, pdflatex get .pdf) 
% To convert them: epstopdf myimage.eps
\begin{document}

%\teaser{
%  \includegraphics[width=\linewidth]{sample.pdf}
%  \caption{Optional image}
%}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

\maketitle

%% Abstract section.

\begin{abstract}
Em jogos {\it massivaly multiplayer online} (MMO) a existência de um mundo virtual persistente é um tópico importante para manter o jogo atrativo e divertido ao jogador. Quanto maior o mundo a ser explorado, tecnicamente mais tempo o usuário passará jogando para conseguir explorar o maior número possível de lugares. A solução proposta neste trabalho é uma ferramenta capaz de gerar mundos virtuais pseudo-infinitos com diversificação de formas e relevos ao longo de sua extensão. Utilizando uma combinação de algorítimos e métodos de gerenciamento de conteúdo, a ferramenta é capaz de criar ilhas/arquipélagos, baías e costas contendo falésias, praias e rochedos que imitam as paisagens encontradas na natureza. Dentre as inovações apresentadas, estão a criação de um terreno virtual de vastas proporções com enfoque mais detalhado no que diz respeito à geração da costa.
\end{abstract}

%% The ``\keywordlist'' command prints out the keywords.
\keywordlist
\contactlist

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introdução}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The computer games market has been evolving considerably over the years. Since the first console, the hardware performance has increased and new graphic technologies were developed, resulting in a wide range of themes and game styles. In the multiplayer games, players interact with other human beings and also with NPCs, which are represented by virtual characters. The kind of game is popular and the social interaction between players is a matter of research ~\cite{onlineGaming,ducheneaut}. In the category of multiplayer games there are the massively multiplayer online (MMO) ones, which are online games featuring large number of players interacting with each other in a persistent virtual world.

An MMO can feature millions of players, such as EverQuest ~\cite{EverQuest} and World of Warcraft ~\cite{Blizzard}, the latter with more than 6 million subscribers ~ \cite{vicioEstrutura}. A persistent virtual world is an important topic to keep the game fun and attractive to the player. The bigger is the world to be explored, technically the bigger is the time the player has to spend in order to explore all the places. As a result of such huge virtual worlds, the creation (and subsequent upgrade) of those worlds is a complex task. EverQuest and World of Warcraft present a virtual world with a wide diversity of geographical features such as mountains, valleys, forests, fields, caves, etc.., and the vast majority of them nominated and related the history of the game. The manual creation of those virtual worlds requires a team able to create heightmaps, adorn landscapes, ensure usability of the map (avoid unreachable places, for example), create interesting place for players, etc. To help on that task, the development of a tool able to generate complex virtual worlds is useful to speed up the development of 3D games such as MMOs.

The solution proposed in this work is a tool able to generate complex virtual worlds in real-time using noise-based techniques of terrain generation. The proposed tool is able to generate complete virtual world featuring continents, oceans, flat areas, mountains, etc, with all those elements generated in real time, on demand as the player moves along the world.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The tool was designed to allow developers to use its features in order to generate 3D terrains for games, particularly MMOs, with minimal human intervention in the generation process. The size of the resulting world is customizable, but it can reach billions of pixels. The use of the term "pseudo" is necessary due physical limitations in computers hardware: an unsigned integer, for instance, can store a certain amount of data; if there were no physical limitations, the tool would be able to generate, in fact, an infinite world.

The content generation is made on demand. As the user moves along the world, the elements inside the user's view are processed and stored into the memory and the ones away from the user's view are removed. Even though the generation of all elements is based on random calculations, if the player visits an specific point {\tt A}, then walks for miles generating a completely different set of landscapes, and returns to point {\tt A}, the same previously seen landscape will be shown again.

Besides the heightmap generation, the tool is also able to create oceans and continents, all of them customizable. Unlike other related works which deal with the generation of continents as a consequence of the heightmap generation, this paper presents as its main contribution a new approach for content generation in virtual worlds. In this approach, the generation of continents, topography and coastlines are handled separately.

The shape of continents is defined on a macro level as a result of an algorithm for maps generation that work projecting pixels onto a sphere. The coast line generation is performed on a micro level. As a result, the tool is able to produce cliffs, beaches, bays and islands.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related works}
\label{sec:related-work}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%As seções seguintes apresentam diversas abordagens utilizadas para a criação de mundos virtuais finitos e pseudo-infinitos, bem como formas de geração de conteúdo para eles (criação de relevo).

\subsection{Virtual city}

The procedural content generation is an old topic in the field of computer graphics. The application of such technique to generate a complete virtual world was used by ~\cite{infinicity}, whose goal was to create a virtual city that was visually interesting and complex.

In the approach the world were divided into a grid consisting of several squares, called cells. The location of each cell is used with a global seed as an input for a hash function ~\cite{wangHash}. The result of this function is used as a seed for a pseudo random number generator and it will define all the characteristics of the buildings within a cell. As a consequence of that approach the contents of a cell is always the same, no matter how far the user moves or if the cell is removed from memory.

To ensure acceptable use of memory and CPU, only the elements inside the user's view are generated. As the user moves through the city, new cells are added to the user's view and the new content is generated. Whe the cell leaves the user's view, it is removed from memory and its resources are released. The cells are placed in square loops around the user and are considered inside the user's view if it is close enough and if it is located within a 120° view range.

\subsection{Layered generation and texturization}

Another related work was a tool used to build the SkyCastle multiplayer game engine~\cite{LandscapeGeneration,SkyCastle}. For the heightmap generation, parameterized procedures and fractal based systems are combined in layered approach: starting with a base map, the application merges a new map with to the base in each iteration. The new maps are pre-calculated and generated using Perlin noise.

One of the texturization methods presented are the use of a large image able to cover the whole world. Although this approach is useful for small scenes, it is not suitable for large scenes nor virtual worlds, since the image size could reach prohibitive proportions. To avoid this problem, an reticulate approach is suggested ~\cite{CSHD03}; using that approacha a texture cell is created so that the a combination of them produces a smooth and continuous and textured plane. The results are acceptable, but it is not visually attractive to the user, because the landscape features an unusual pattern that will never exist in the real world. In order to achieve a better visual result, an approach proposed by ~\cite{LN03} is suggested. In that approach, a set of pre-defined border textures is combined with reticulate textures. The idea is to apply the border textures on an already textured plane, but using transparency in the border textures. As a consequene, border textures will overlap the some reagions of the plane and produce a less homogenous landscape, which creates a better visual result.

In order to decorate the virtual world, three procedural plant generation methods are used: L-System ~\cite{PL90}, component-based generation ~\cite{LD98} and parameterized trees ~\cite{WP95}.

\subsection{Procedural planets}

The use of fractals and their combinations with other methods are also used to proceduraly generate planets ~\cite{LindaOndrej2007}. In this approach, the generated world is not infinite, but it is spherical and simulates the view of the planet Earth. Starting with a recursive subdivision of an octahedron, a spherical world is created and it is used as a first step to sevel algorithms of heighmap generation. The heightmap is obtained as a result of several techniques, each one featuringits peculiarities and results. Among the techniques used are: generation using random failures, midpoint displacement (including a multifractal variation) and Perlin noise (and its many variations).

\subsection{Erosion fractals}

Another approach for procedural terrain generation in real time uses fractals affected by erosion ~\cite{terrain_generation}. As pointed out, the increasing processing power of home computers has enabled games to make use erosion simulations almost in real time. As a consequence, an approach envolving fractals and erosion is possible to be reached, resulting in more realistic terrain. In order to achieve that, a heightmap was created from a Voronoi diagram, where each vertices of the diagram is called {\it functionality point}; the value of each cell in the heightmap is obtained through a linear combination of distances of the closest functionality points. Once the heightmap is generated, it is combined with a noise algorithm in order to produce mountains. The sharp lines introduced by the well-defined Voronoi diagram are removed with the application of a turbulence filter. For the erosion simulation an hybrid method is used, which is a combination of two techniques: the thermal ~\cite{MusgraveErosao} and the hydraulic erosion.

\subsection{Stochastic subdivision of a quadtree}

Another related work uses stochastic subdivision techniques in order to generate a virtual world ~\cite{dollins-thesis}. The main goal was to create a virtual world of large proportions, however the creation of its content is procedurally made and performed on-demand with a multi-resolution approach. For the heightmap generation a recursive sub-division process is applied to a quadtree. The height of each vertex is defined by the midpoint of each cell in the quadtree. For each level of subdivision, the midpoint of the each new child cell is defined by the parent cells around the cell being divided. The height of all other vertexes, such as the one at the corners, is defined as a result of an interpolation of the closest midpoints. In each level of detail, more cells are divided and replaced with the new generated child cells.

\subsection{Map generator}
\label{sec:geradortorben}

Another related work focusing on the generation of maps for planets, which is based on the generation of spheric worlds using a recursive subdivision of a tetrahedron ~\cite{torben}. Although the work is not presented as a graphical tool where the user can explore a 3D environment, all generated information is part of a complete virtual world featuring highly customizable continents and oceans. The presented idea is to create a map as a result of a projection of pixels onto a sphere, a method similar to ray tracing~\cite{raytracing}. In the same way it is performed in ray tracing, where rays leaving the camera collide with something and create a visible pixel, each pixel of the map being generated is projected onto sphere. The projection algorithm sets a height value to each pixel found, resulting in a complete heightmap.
In order to render the map, initially the visible points on the sphere surface are found. After that, the height value for each one of those points is calculated as follows:

\begin{itemize} 
  \item Insert the sphere inside the tetrahedron.
  \item Divide tetrahedron into two smaller tetrahedral. 
  \item Choose which of the two tetrahedral contains the point being analyzed. 
  \item Repeat the above steps until the tetrahedron is small enough.
  \item Use the average height of the each tetrahedron vertex in order to calculate the height of the point being analyzed. 
\end{itemize}

Initially the sphere is placed inside a irregular tetrahedron and information about height and a random seed are set for each of the vertexes. After the tetrahedron is divided in two pieces by a plane formed between the midpoint of the longest edge and the two end points of the edge in front of the midpoint of the longest edge. In short, each of the two resulting tetrahedral have three points from the original tetrahedron and a new vertex placed in the midpoint of the largest edge. The height and the seed of the new vertex is calculated from the height and seed of the end points of longest edge. The process is repeated recursively for each of the tetrahedral until it is small enough to contain only the desired point.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tool organization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The following sections explain in a top-down approach how the tool was structured to achieve the described goals.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Initial analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

There are several related works concerning the generation of virtual worlds, whether finite or infinite. One of the realted works creates an infinite city ~\cite{infinicity} that is presented to the user on demand as it walks on the ground. This work was the ground zero for Charack development, however the original idea was changed in order to make the tool suitable to generate more types of terrains (mountains, plains, continents, etc.), not only streets and buildings. The approach of content generation made on demand was maintained.

The procedurally generated world approach ~\cite{LindaOndrej2007} is very close to the concept of content generation aimed for Charack. In that work, a spherical planet is created as a result of a recursive division of a geometric shape, then noise functions are applied to the mesh to generate the heightmap. There is no distinction between the content generation approache for continents and the content generation for the lands within the continents. As a result the continents are created by flooding the heightmap with a water plane, which will produce the coast lines based on height of the sea level and the amount of ripples in the topography. The content itself is not generated on demand. Charack was created from an evolution of this idea, but subject to some limitations. The world created by Charack handles differently the generation of continents, coast lines and the heighmap within them. Each of these generation processes has its own peculiarities, which can be tweaked in order to customize specific details in the world with no interference in the other processes. This is the case of the creation of long beach on the coast without changing the basic structure or the topography of the continent that the beach is inserted. Charack also generates the content on demand, however it doesn't use a spherical approach, the resulting world contained within a plan.

The generation of a virtual world as a result of recursive subdivisions of a quadtree ~\cite{dollins-thesis} is very similar to the Charack proposal. In that work, a world with huge proportions is created and its content is generated on demand as the user moves. The heighmap is created in a parameterized and multi-resolution way, so the closer the user of place, the greater is the amount of detail there. There is also no distinction in the generation process of continents/coastlines/land content. The heightmap generation proposed is used by Charack, however continents and the coastlines generation process are completely different.

Analyzing the related work, virtual worlds are generated several approaches, but in none of them handles differently the content generation for continents, coastlines and topography. Although there are variations in how the heightmap is created, the generation of continents is a result of a water flooding plane. This approach allows the developers to focus on the content generation for the land, however it has a simple approach concerning continents and coastlines. The main idea and contribution of Charack is the content generation handled differently for each world element (continent, coastline, etc.), with an aggressive and specific approaches for each one. Most of the related work focuses on heightmap generation, so we believe the Charack contribution will be better if the efforts are focused on the generation of continents and coastlines.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Basic structure}
\label{sec:basic-structure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In order to create a virtual world that reaches the presented proposal, a top-down approach is used for the content generation. The Charack data flow begins in a macro view of the world, which are the continents, evolving to a micro view of the planet, which are the generation of content for each vertex that will be drawn in the screen. Figure ~\ref{fig:planejamento_estrutura_basica} shows Charack basic structure.

\begin{figure}
\centering
\resizebox{5cm}{!}{\includegraphics{images/planejamento_estrutura_basica.png}}
\caption{Basic structure for content generation}
\label{fig:planejamento_estrutura_basica}
\end{figure}

\subsubsection{Maps generator}

At the top of the chain is the maps generator, which creates the continents that exist throughout the virtual world. This module is an encapsulation of the solution created by ~\cite{torben}, which was described in the section ~\ref{sec:geradortorben}. When the tool is initiated, it uses a user defined seed to generate all the continents. Once the continents are generated, all the information related to terrain types (land, water and coast) are stored in a matrix, called {\bf macro-matrix (MM)}, which is used by all the other algorithms.

\subsubsection{Slice manager}

Below the MM and the maps generator is the slice manager (SM). It extracts a portion of the virtual world (the user's view) and provide the render engine with information about the heightmap. In order to obtain the required information to create the heightmap, the slice manager uses the coastline generator (CG), which uses height generator (HG) and the data stored in the MM.

In the context of the SM, there is no information about land or water, it only knows a set of pixels of the virtual world and its height values. Using the position of the user as a guide, the SM slices virtual world and, for each collected pixels, it queries the CG in order to find out its height value.

\subsubsection{Height generator}

The height generator (HG) defines the height value for each pixel in the virtual world. To ensure that the developers can create a customizable heightmap based on their needs, new functions to generate content can be added to the tool in a simple way.

\subsubsection{Coastline generator}

The coastline generator (CG) will map each pixel of the slice manager to the MM in order to find out the terrain type of that pixel. If the pixel being analyzed is mapped to a location in the MM that is described as water, then the CG assigns a height value equals to sea level for the pixel and returns it to the SM. If the pixel is mapped to a place described as (simple) land, then the CG will use the information provided by the HG in order to find out the height value for that pixel. Finally, if the pixels is mapped to a place described as coast, then the CG uses its own structure (together with the MM) to set height value for the that pixel.

The resulting virtual world technically has height and width defined by the maximum size of a signed integer. It is physically impossible to generate a MM with such proportions. Since the MM is smaller than the virtual world, an MM's entry {\tt (i, j)} represents sevel pixels in the virtual world. Figure \ref{fig:mapeamento_macro} illustrates the MM mapping process.

\begin{figure}
\centering
\resizebox{4cm}{!}{\includegraphics{images/mapeamento_macro.png}}
\caption{Mapping the MM to the virtual world: each MM pixel represents several pixels in the virtual world.}
\label{fig:mapeamento_macro}
\end{figure}

As the figure ilustrates the smaller is the MM, the more pixels in the virtual world will be represented by the same entry in the MM. If the virtual world had {\tt 1000x1000} as its size and the MM had {\tt 10x10} as its size, for instance, it means that for each pixel of the MM represents 100 pixels in the virtual world. If one of these pixels is described in the MM as coast, then there is an area of {\tt 100x100} pixels in the virtual world that must be a coast. To solve this problem the CG acts. When any of those pixels in that particular area of the virtual world is analyzed by the CG, it will work on it and return it with different values, which will result in a coast line for that area, not am area entirely filled withby land or water.

\subsubsection{Rendering engine}

The rendering engine draws created heightmap in the screen. The result is rendered as triangles mesh that is textured according to the height value of each vertex in the mesh.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation}
\label{sec:implementation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The main problem concerning Charack's implementation was the on demand content generation. Based on the fact the user can only see what is inside the view area, all the content generation algorithms need to take into account {\it only} the information that is available in the user's view. Even though this approach is efficient for resources management (process only the visible elements), it increases the complexity of the content generation algorithms.

The algorithm that generates mountains, for instance, has no way to determine where the mountain ends, because the world outside the user's view technically does not exist yet, it will be generated as the user moves. One approach to solve that problem would be the use of a function that describes the mountain chain, but this function should not rely on begin/end points, because they could not exist in a certain time. If that function does not need any begin/end points, at least it would have to rely on the position of the user in the virtual world. If the function must be aware of some special points, those points have to be previously processed, which would break the on demand content generation concept.

In addition the algorithms are drastically affected by the fact that the information they receive in a certain time may disappear altogether in the next iteration, since the user can move and change the visible content. Using the example of the mountain generation, a mountain could present an abrupt end, because the points being used for the content generation left the user's view.

To circumvent these problems the content generation was divided into three main stages: infinite terrain, continents and height generation. The on demand content generation affects differently each of these stages and the problems and solutions related to each stage are described in the following sections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Infinite terrain}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The fundamental idea for the content generation in the virtual world is the possibility the user has to walk in a infinite way, looking at new contents as the movements are performed. As the user walks, the tool must be able to identify where the observer is located in the world in order to generate the content around that position. To solve this problem, a variation of a technique described by ~\cite{infinicity} was used.

The solution makes the player able to look at the screen and see a slice of the virtual world, but with no explicit divisions in the world, such as cells. Unlike what was done in ~\cite{infinicity} where the user's view is a cone, the Charack user's view is a square centered on the user. Using the user position {\tt (x, y, z)}, the visible content arround that coordinate is sliced from the world and drawn in screen. If the player reaches a place or a complete slice is not possible, such as the world boundaries, no content is displayed beyond the world area.

In order to texture the sliced data a set of images are interpolated and managed by the shading language GLSL ~\cite{glsl}. The height value of the pixel defines the interpolation weight of each texture. As a consequence, a sand texture has highter weight for pixels featuring a low height value, for instance. Figure ~\ref{fig:set_texturas} shows all the available textures.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/set_texturas.png}}
\caption{Set of image used for terrain texturization}
\label{fig:set_texturas}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Height generation}
\label{sub:relevo}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The main idea for the terrain height generation is the use a parametric function that informs the height value of each vertex. The function is seeded with the point location in the world. As a consequence, the function is able to describe all the height information in the world with no limitions concerning the world size. The processing time is related to the size of the user's view, not related to the world size, since the function uses the point information to calculate its height value. The height values are generated with a Perlin noise function ~\cite{Perlin85}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Continents}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The generation of continents and oceans has been proposed in order to break the monotony of a landscape composed only by land and to increase the similarity of the virtual world with the real world landscapes. The solution for the continents generation consist in pre-process the land areas and store that information for other tool calculations. With that approach, the on demand content generation has been partially broken, since the continents are generated before all the other content, but it ensures a better control over water/land areas.

The continents generation is based on the approach described in section ~\ref{sec:geradortorben}. That planet generator was used because it has several parameterization options, such as the possibility to use a seed to manage all the random calculations. Figure ~\ref{fig:torben} {\it (a)}, {\it (b)} and {\it (c)} illustrates the results obtained with the planet generator of  ~\cite{torben}.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/torben.png}}
\caption{Random planets generated with different seeds. (a) and (b) maps featuring height value information; (c) map featuring only information about what is wahter/ocean.}
\label{fig:torben}
\end{figure}
%~\cite{torben}

\subsubsection{Problems with continent generation}
\label{sec:visaomicromacro}

As previously explained in section ~\ref{sec:basic-structure}, each pixel of the MM is mapped to several pixels in the virtual world. A direct consequence of that mapping process is the generation of large areas featuring straight land lines. If there were no hardware limitation and if it were possible the generation a matrix with the exact size of the virtual world, the matrix would contain the necessary resolution for the tool to accurately determine whether a pixel is land or not land, in a ratio of 1:1 (one MM pixel is mapped to one world pixel). This approach, however, is not suitable because a matrix with such proportions consumes many resources and processing time. Although the tool allows customization of the MM size, tests presented that a {\tt 800x800} pixels MM has is quite enough information to be processed by all the other algorithms of the tool.

Figure \ref{fig:ilustracao_mapeamento_macro} illustrates the results obtained by the tool when no algorithm is used to generate extra content to fill the rectilinear spaces of the virtual world.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/ilustracao_mapeamento_macro.png}}
\caption{The result of no algorithm to generate extra content to fill the discrepancies in the MM mapping process}
\label{fig:ilustracao_mapeamento_macro}
\end{figure}

This figure ilustrates a place in the virtual world that represents the transition between two different points of the MM (a land point and a water point). To explain what is happening, assume the tool is drawing the world at position {\tt (x, y, z)}, which is the mapping result of a pixel {\tt (i, j)} in the MM, which is described as land; as the tool increases the coordinate in order to draw the landscape, each new position is mapped to the MM. If the result of the mapping process of new coordinate, {\tt (x + 1, y, z))} for instance, is still the point {\tt (i, j)} in the MM, then the tool will again draw a land pixel on the screen. Assuming that only at point {\tt (x + 10, y, z)} the pixels start being mapped to a different pixel in the MM, such as {\tt (i + 1, j)} (and {\tt (i + 1, j)} is described as land), then all points before the position {\tt (x + 10, y, z)} are drawn as water and all pixels after that location are drawn as land.

The figure shows clearly when the world coordinates start beeing mapped to a different entry in the MM, which is when the tool replaces the lang rendering with water rendering. As a consequence of no algorithm being applied to generate content for that transition area, the user will move along the coastline and will see only straight lines.

\subsubsection{Coastline disturbance}
\label{sec:coastline-disturbance}

The mapping process of the pixels of the virtual world to the MM produces very unrealistic landscapes. A real world beach has a natural curvature and hardly have a length of 20km in a perfectly straight configuration, as the beaches generated by Charack. Although the objective of this work is not to create photorealistic landscapes, such unreal beaches are not acceptable. To circumvent this problem, a coastline disturbance is applied to the locations where the mapping process is made between two MM points, one of them described as land and the other one described as water. The algorithm is described below.

The MM has a full description of what is land and what is water in the virtual world. Each of its pixels has a descriptor, which tells the other algorithms what type of terrain one pixel of the virtual world is after it is mapped to the MM. Charack features three types of terrain: water, land (continent) and offshore (land in contact with water). After the continents are pre-processed and stored in the MM, it only features information about land (continents) and water.

From that moment, the first step of the coastline disturbance algorithm is performed. Using the current MM as its input, the algorithm scan each MM's pixel, updating the descriptor of pixels that represent a coast. A pixel is said to be coast when at least one of its neighbors is water. After the algorithm end, the MM contains the three types of terrain described before (water, land and offshore). The next step to apply disturbance to the coastline is the content generation based on the descriptor of each pixel in the MM. When the tool is creating content to draw on the screen, each pixel being drawn is tested against its descriptir in the MM. If the pixel is mapped to a land pixel in the MM, then the function will set a height value for that point. If the pixel is mapped to a water pixel in the MM, then the function will set the sea level height to that point. Finally if the pixel is mapped to a offshore pixel in the MM, then the function will disturb the land/water information of that mapping process, which will result on a non-straight coast line. Figure ~\ref{fig:funcionamento_quebra_linearidade} illustrates the algorithm.

\begin{figure}
\centering
\resizebox{6cm}{!}{\includegraphics{images/funcionamento_quebra_linearidade.png}}
\caption{Coastline disturbance algorithm}
\label{fig:funcionamento_quebra_linearidade}
\end{figure}

The MM pixels {\tt A} and {\tt B} have a descriptor indicating that they are described as a coast. Some other pixels of the figure are also coast, but they will not be detailed for understanding purposes. Plan {\tt M} describes the MM and plan {\tt V} describes the result of the mapping process between them. It is not described in the figure, however each blocks of plan {\tt V} is composed of several pixels, while each block of plan {\tt M} represents only one MM pixel. The MM pixel {\tt C} is mapped to a massive block of land in the plan {\tt V}, as its descriptor tells the tool that it is a pixel described as land. The pixel {\tt A} would also be mapped to a massive block of land, but with the intervention of the coastline disturbance algorithm it is mapped to a different configuration. During the content generation for the pixels that are inside the block {\tt RA}, the coastline disturbance algorithm alters the land/sea information for each pixel, so that the block will not be composed of land or water pixels only, but a combination of them instead.

The implementation of that process is based on a noise function and random numbers with a parametric function deciding what is land and what is water for all pixels described as a coast in the MM. Using the pixel position in the block {\tt RA}, the function maps that information into a spectrum of values created by a Perlin noise function. What the parametric function does is check if the hash of the pixel being analyzed is inside or outside of the spectrum. The process can be imagined as a height test against a small heightmap (which is created as result of the noise spectrum): if the return of the noise function for that is greater than a certain value (which is the granularity of the block being analyzed), then it is classified as land, otherwise it is classified as water. The higher is the granularity of the block, the greater is the amount of land on that location. Figure ~\ref{fig:plot_ruido_quebra_linearidade} illustrates the small heightmap generated by the costline disturbance algorithm when block {\tt RA} is being processed.

\begin{figure}
\centering
\resizebox{2cm}{!}{\includegraphics{images/plot_ruido_quebra_linearidade.png}}
\caption{Small heightmap generated by the costline disturbance algorithm}
\label{fig:plot_ruido_quebra_linearidade}
\end{figure}

\subsubsection{Beaches}

The coastline disturbance algorithm minimizes the problem of unrealistic continents lines, but the outcome is not quite good enought. When Charack is rendering a slice of the world, for each pixel described as a land a height vaule is set to the pixel; the same applies to the pixels that are described as water, but in that case the height value is always the same (the sea level). As a direct result of that approach if the tool is drawing a set of pixels describing a mountan and the next pixels are described as water, the landscape will features a "step". It happens because the mountain chain was generated very close to the water, which means that its rendering is abruptly interrupted when Charack finds pixels described as water. Although there are cliffs in the real world, they are not present in all coasts. To solve this problem, a special algorithm is applied in order to create beachs in certain locations of the world, which makes the generated landscape looks more realistic.

The algorithm for beach generation is performed right before the content is rendered on the screen. After Charack maps the pixels to the MM and after the coastline disturbance algorithm is performed, the result is a heightmap ready to be rendered. Before being drawn on the screen, the heightmap is treated by the beach creator algorithm. The procedure scans each pixel in the map and for each one it checks the distance that the current pixel is from a water pixel around them. The pixels around the target are mapped directly to the MM, so the only information that is used from the heightmap is the pixel location in the world (which is necessary to map it to the MM). The checking process is performed in four directions (right, left, up and down) and the tool keeps on searching until it finds a water pixel or when {\tt N} pixels are analyzed. After that, the four distances are added and used to calculate the height of the beach. The possible results are:

 \begin{itemize}
 \item Se o pixel analisado estiver à uma distância de {\tt 4N}, isso quer dizer que a ferramenta percorreu às quatro direções possíveis e não encontrou água. Nesse caso, o pixel em questão não tem a sua altura recalculada; esse caso descreve o que acontece com todos os pixels que estão dentro do continente ou na costa porém longe da água: eles não formam uma praia e sua altura é definida pela função de relevo principal;

 \item Se o pixel analisado estiver à uma distância inferior a {\tt 4N}, então a sua altura será recalculada. Quanto maior for a distância calculada, maior será a altura do pixel, porém essa variação da altura é calculada dentro de um intervalo definido {\tt [T,B]}, onde {\tt T} é a altura máxima e {\tt B} é a altura mínima de uma praia, respectivamente. O resultado dessa abordagem é uma praia que inicia alta no continente e, à medida que se aproxima da água, fica mais baixa.
 \end{itemize}

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/exemplo_praia_simples.png}}
\caption{Praia gerada ao longo da costa}
\label{fig:exemplo_praia_simples}
\end{figure}

A figura ~\ref{fig:exemplo_praia_simples} ilustra o resultado obtido com a geração de prais.

\subsubsection{Arquipélagos e praias diferenciadas}

Utilizando os algoritmos de quebra de linearidade e criação de praias ao longo da costa, a ferramenta passou a gerar paisagens mais realistas. O resultado final em relação à costa e à praia, porém, apresentou um padrão muito definido, o que é incomum de acontecer no mundo real, no qual as linhas e paisagens naturais tendem a seguir um princípio aleatório ou menos padronizado. Se o usuário viajasse pelo mundo virtual apenas pela costa, ele veria prais com a mesma configuração (mesmo tamanho) e nenhuma ilha ou arquipélago ao longo do oceano. Para melhorar esse aspecto, criaram-se dois novos algoritmos que atuam na costa: um diferenciador de prais e um gerador de arquipélagos.

O \textbf{diferenciador de prais} atua perturbando a distância utilizada para o cálculo dos pixels água vizinhos a um determinado pixel. Em vez de utilizar uma distância fixa {\tt N} para o cálculo da distância até a água, o diferenciador utiliza a posição do pixel como semente para uma função de ruído, que tem como retorno a nova distância que será utilizada nos cálculos. Utilizando essa técnica, o diferenciador é capaz de alterar o tamanho e forma da praia, o que faz com que determinadas regiões apresentem uma maior quantidade de areia do que outras. A figura ~\ref{fig:exemplo_praia_complexa} ilustra os resultados obtidos.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/exemplo_praia_complexa.png}}
\caption{Resultado obtido com a aplicação do diferenciador de praias}
\label{fig:exemplo_praia_complexa}
\end{figure}

O \textbf{gerador de arquipélagos} atua criando novas faixas de terra em terminados pixels da MM. Depois que a MM é criada e todos os descritores de informação de cada um de seus pixels é configurado, o gerador de arquipélago itera sobre os pixels que representam a costa e, para alguns deles, adiciona a informação que essa região possui ilhas. No momento em que a ferramenta estiver renderizando os pixels que são mapeados para essa região da MM, o descritor de informação será consultado e a ferramenta saberá que essa região necessita de um conteúdo novo, além do conteúdo utilizado para a quebra de linearidade da costa. Esse conteúdo é criado utilizando a mesma abordagem do algoritmo de quebra de linearidade, porém utilizando como pixels de análise aqueles pixels que são água. Para cada pixel sendo analisado, a sua posição é utilizada como um hash que é testado contra um espectro criado por uma função de ruído. Os ruídos utilizados para esse espectro são diferentes daqueles do algoritmo de quebra de linearidade, visto que o resultado esperado são pequenas porções de terra, não um bloco maciço dela. Para conseguir esse efeito, aumentou-se o número de oitavas da função de ruído de Perlin e aumentou-se o limite que é utilizado nos testes para saber se o pixel é ou não terra. O resultado do gerador de arquipélagos é combinado com o algoritmo de quebra de linearidade, o que faz com que as ilhas sejam geradas, em determinados casos, muito próximas à costa. O resultado final obtido com o gerador de arquipélagos são ilhas de diversos tamanhos ao longo da costa em determinados locais do mundo virtual. A figura ~\ref{fig:exemplo_ilha} ilustra os resultados obtidos.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/exemplo_ilha.png}}
\caption{Ilha obtida com o gerador de ilhas}
\label{fig:exemplo_ilha}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Resultados}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Essa seção tem por objetivo avaliar cada uma das técnicas utilizadas na geração de conteúdo da ferramenta, mostrando os resultados que foram obtidos com cada abordagem. Leva-se em consideração a flexibilidade proporcionada pela ferramenta durante a utilização da funcionalidade e o resultado gráfico alcançado.

É importante frisar que o objetivo da ferramenta, no escopo do presente trabalho, não é a geração de conteúdos reais ou foto-realísticos, mas sim de elementos que possam ser utilizados para a criação de um cenário num jogo 3D. Entende-se por graficamente aceitável todo o resultado obtido que se enquadre dentro de um jogo e não destoe daquilo esperado pelo jogador, como uma montanha composta por ondulações senoidais suaves ao invés de um conjunto de cristas piramidais.

\subsection{Avaliação dos continentes}

O tempo para a geração dos continentes é diretamente proporcional ao tamanho da macro-matriz especificada. Em testes realizados, a redução da macro-matriz para valores inferiores a {\tt 800x800} rendeu aumentos de desempenho consideráveis. Em contra partida, quanto menor o tamanho da macro-matriz, mais quadrados e lineares serão as linhas da costa de cada um dos continentes, o que pode produzir um resultado gráficos não muito aceitável. Para contornar esse problema, é possível ajustar o algorítimo de geração da costa para que ele faça alterações mais agressivas nas bordas dos continentes, o que irá quebrar a linearidade gerada por uma macro-matriz de baixa resolução. O ajuste desses dois elementos abre um leque de possibilidades para o programador, que pode encontrar um ponto de equilíbrio entre tempo de geração dos continentes e linearidade da costa. A figura ~\ref{fig:charack_final_continentes} mostra os resultados obtidos em relação a continentes e oceanos gerados pela ferramenta.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_continentes.png}}
\caption{Continentes e oceanos gerados pela ferramenta}
\label{fig:charack_final_continentes}
\end{figure}

\subsection{Avaliação do relevo}

O relevo gerado pela ferramenta é inteiramente parametrizável e customizável pelo programador. Seguindo o protótipo da função esperada pela ferramenta para o cálculo do relevo, o programador pode criar qualquer código para a geração de relevo, o que garante flexibilidade nesse aspecto. A ferramenta possui embutida um gerador de relevo baseado na função de ruído de Perlin, que produz uma paisagem semelhante àquela encontrada no mundo real. A figura ~\ref{fig:charack_final_relevo} ilustra o relevo criado pela ferramenta utilizando-se as funções de geração embutidas.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_relevo.png}}
\caption{Relevo gerado pela ferramenta}
\label{fig:charack_final_relevo}
\end{figure}

A função de relevo que foi implementada na ferramenta é capaz de gerar paisagem que imitam de forma aceitável o que é encontrado na natureza, porém a grande maioria dos esforços de desenvolvimento não foram empregadas sobre esse tópico. Como consequência, a ferramenta não possui um grande leque de possibilidades de geração de conteúdos para o interior dos continentes, como cadeias montanhosas, planícies e planaltos. O relevo gerado atualmente é simplista no sentido de não apresentar grandes diversidades geológicas, entretanto se o objetivo da utilização da ferramenta não for criar um terreno rico em diversidade, o objetivo é plenamente atingido.

A geração atual de relevo foi desenvolvida visando-se a obtenção de ondulações de média/baixa frequência nas funções de ruído, o que faz com que não existam cadeias montanhosas pontudas ou depressões abruptas. A baixa frequência nos ruídos da função de geração de relevo faz com que o usuário veja montanhas com um ângulo muito suave de inclinação, além de áreas com poucas ondulações, que podem ser interpretadas como planaltos. A figura ~\ref{fig:charack_final_relevo_plano} ilustra uma área do mundo virtual que possui relevo com baixa frequência.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_relevo_plano.png}}
\caption{Área com relevo de baixa frequência}
\label{fig:charack_final_relevo_plano}
\end{figure}

Se a função de geração de relevo utilizar frequências muito baixas e o seu espectro de valores for estendido à todo o mundo virtual, as montanhas e ondulações que serão produzidas serão muito suaves. Esse fenômeno acontece porque baixas frequências não criam alterações acentuadas de altura nos polígonos do relevo. Na utilização de altas frequências, a grande quantidade de ruído cria montanhas mais pontiagudas, porém as áreas intermediárias entre elas são muito ondulada, que é o reflexo da propagação de um esprectro de valores muito ruidoso para o mundo virtual. Uma das formas de contornar esse problema é replicar a extensão do espectro de valores, ou seja, em vez de utilizar um espectro que cubra o mundo inteiro, utilizar esse mesmo espectro três vezes ao londo do mundo. Isso permite que pouco ruído seja utilizado, porém evita que o relevo resultante tenda ao plano. A ferramenta utiliza essa abordagem para criar um relevo com montanhas consideravelmente onduladas, porém sem áreas de ondulação estranha entre elas. A figura ~\ref{fig:charack_final_relevo_comparacao} ilustra os diferentes tipos de relevo obtidos com a variação do tamanho do espectro de valores.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_relevo_comparacao.png}}
\caption{Diversos relevos gerados a partir de diferentes espectros de ruídos. (a) Pouco ruído estendido para o mundo inteiro sem replicação. (b) Pouco ruído replicado 200 vezes. (c) Bastante ruído com replicação de 200 vezes. (d) Bastante ruído sem replicação.}
\label{fig:charack_final_relevo_comparacao}
\end{figure}

\subsection{Avaliação da costa}

A geração da costa é composta por dois pilares principais, um de aspecto mais global e outro mais local. No aspecto global, a ferramenta utiliza apenas dados encontrados na MM para criar as linhas que compõem a costa, conforme descrito na seção ~\ref{sec:visaomicromacro}. O resultado final para essa abordagem são costas completamente retilíneas, o que é irreal do ponto de vista do usuário. A figura ~\ref{fig:charack_final_costa_reta} ilustra duas linhas da costa completamente retilíneas.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_costa_reta.png}}
\caption{Local de encontro de duas linhas da costa sem a aplicação de qualquer algorítimo de geração de conteúdo extra}
\label{fig:charack_final_costa_reta}
\end{figure}

Depois que o método para quebra de linearidade da costa foi implantada, a ferramenta passou a apresentar paisagens mais aceitáveis. As figuras ~\ref{fig:charack_final_costa_1}, ~\ref{fig:charack_final_costa_2} e ~\ref{fig:charack_final_costa_5} ilustram a criação de pequenas baías em certos locais da costa resultantes da aplicação do algoritmo. Isso acontece porque nesses locais o algoritmo de quebra de linearidade da costa criou braços de terra curvos partindo da linha reta do continente e, ao mesmo tempo, o algorítimo de criação de praias reduziu ao máximo a quantidade de areia encontrada no local. A ferramenta também é capaz de criar golfos, que são baías de grandes proporções, porém não é possível prever o local exato que isso irá acontecer, porque tal resultado depende da combinação de valores e coordenadas que variam sensivelmente ao longo do mundo virtual.

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_costa_1.png}}
\caption{Pequena baía com rochas}
\label{fig:charack_final_costa_1}
\end{figure}

As figuras ~\ref{fig:charack_final_costa_3} e ~\ref{fig:charack_final_costa_4} ilustram a adição de conteúdo à costa retilínea do continente, resultando numa paisagem mais convincente para o usuário. As duas figuras mostram com bastante detalhe a combinação da função de geração de relevo, o diversificador de praias e o algorítimo de quebra de linearidade da costa na criação de conteúdo; a figura ~\ref{fig:charack_final_costa_3} apresenta uma rocha imediatamente à esquerda da falésia principal, resultado obtido a partir de um espectro de ruído que gerou dois pontos principais indicando a existência de terra: o maior sendo a falésia e o menor sendo a rocha grande à esquerda. Além disso, o diversificador de praias removeu a areia na parte de baixo da falésia, porém à esquerda ele fez o processo contrário. Em conjunto, o algoritmo de geração de relevo com replicação de valores criou pontas na extremidade esquerda da falésia.

A ~\ref{fig:charack_final_costa_4} apresenta a mesma combinação do algoritmo da figura anterior, porém o resultado obtido variou em função da coordenada do mundo virtual no local. Ao contrário do que aconteceu anteriormente, o diversificador de praias adicionou areia à base da falésia e, à esquerda da imagem, criou uma extensão da praia em forma de "braço". Também à esquerda, o diversificador de praias removeu a areia que fica na base do continente, criando um aspecto menos padronizado de conteúdo. À direita, é possível observar que o algoritmo de geração de relevo criou um declive que termina de forma relativamente suave na praia (relevo de baixa frequência), ao passo que na parte esquerda da imagem o relevo termina de uma forma mais abrupta (relevo de alta frequência).

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_costa_2.png}}
\caption{Baía de médio porte gerada a partir da criação de um braço de terra originada no continente}
\label{fig:charack_final_costa_2}
\end{figure}

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_costa_3.png}}
\caption{Extremidade de um continente}
\label{fig:charack_final_costa_3}
\end{figure}

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_costa_4.png}}
\caption{Costa de um continente com praias de tamanho variável}
\label{fig:charack_final_costa_4}
\end{figure}

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_costa_5.png}}
\caption{Baía criada a partir da geração de dois braços de terra originados no continente}
\label{fig:charack_final_costa_5}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusão}
\label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A criação automatizada de mundos virtuais procedimentais é uma das formas existentes para auxiliar desenvolvedores de jogos a criarem ambientes mais ricos em detalhes, em menos tempo e com menos recursos humanos. Ao contrário da abordagem puramente não automatizada, no qual um {\it game designer} ou artista deve desenhar e modelar o mundo virtual por completo, na abordagem automatizada a adição de detalhes e modelagem de locais fica a cargo da ferramenta geradora. Dependendo do grau de realismo e da complexidade dos algorítmos de geração de conteúdo, uma ferramenta automatizada pode ser capaz de gerar um mundo virtual aceitável para diversos escopos de projeto.

Existem diversas pesquisas em relação à essa área, com abordagens diferentes e focadas em resultados diferenciados. Pode-se destacar nelas a preocupação em criar um mundo virtual de proporções infinitas sem que isso comprometa o desempenho da aplicação. Para atingir esses resultados, diversas técnicas são utilizadas, como aplicação de LOD e {quadtrees} para gerenciamento de malhas, além de diversos algorítimos procedimentais para a geração de conteúdo, como criação de relevo a partir de fractais, multi-fractais e/ou combinação de funções de ruído.

Este trabalhou apresentou o desenvolvimento de uma ferramenta capaz de gerar mundos virtuais pseudo-infinitos com diversificação de formas e relevos ao longo de sua extensão. Utilizando uma combinação de algorítimos e métodos de gerenciamento de conteúdo, a ferramenta é capaz de criar praias, ilhas/arquipélagos, baías e costas que imitam as paisagens encontradas na natureza. Além disso, a possibilidade de parametrizar cada um desses elementos dá ao desenvolvedor um controle maior sobre o resultado que será obtido.

Dentre as inovações apresentadas, estão a criação de um terreno virtual de vastas proporções com enfoque mais detalhado no que diz respeito à geração da costa. Fruto de um aprofundamento nas pesquisas já realizadas na área, o desenvolvimento da ferramenta foi focado em tratar de forma diferenciada a criação de conteúdo para os diversos elementos existentes (como continentes, relevo, etc). Uma das prioridades do trabalho foi a criação de bordas dos continentes, não a criação de conteúdo para o seu interior; esse relevo está aceitável, porém ainda está muito aleatório e sem grandes resultaods. A ferramenta é capaz de gerar continentes contendo falésias, praias e rochedos em suas extremidades. Em testes realizados, o tempo que um jogador levaria para chegar de uma ponta do mundo virtual até a outra num ambiente com o tamanho máximo suportado por um inteiro sem sinal, andando 100 pixel por segundo, seria de um 1 ano e 3 meses.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section*{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%To Robert, for all the bagels.

\bibliographystyle{sbgames}
\bibliography{paper}
\end{document}